---
title: "Analysis of Zillow Data from Austin, Texas"
author: "Will Curkan"
date: "2022-12-16"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo = F, warning=F, message=F, include=F}
library(dplyr, quietly = T)
library(ggplot2, quietly = T)
```

# Introduction

Austin, a hilly city due south in the middle of Texas, is where all the young, hip Texans and out-of-staters are converging to live. Other than its famous $6^{th}$ street nightlife, the "weird" city is home to many "corporate and regional headquarters" of companies like Dell, IBM, Facebook, VISA and many more [1] - these are some great reasons to "kick up the roots" and mosey on over to Austin!
Suppose you think that you are *just* young and hip enough to move out to Austin, you might want some insights on how the housing market is.

## Objective Statement

The objective of this project is to explore the dataset, make inferences about true statistic parameters in Austin, Texas, and to make predictions on housing prices and categories like "which broker is listing the property".

## Data Acquisition and Explanation

```{r load data, echo = F}
set.seed(2022)
data <- read.csv('Zillow_Austin_11-16-22.csv')
#head(data)
#dim(data)
```
The user Alex Huggler on Kaggle.com provides Zillow Housing data for November 2022 on [2]. The data has 800 unique observations and 20 features: one observation per unique housing unit and each feature as some trait of the house like its zip code or number of bedrooms.

### Feature Descriptions

The following features will be used in analysis of Austin housing prices:

|Feature|Description|
|-------|-----------|
|price| Price of housing|
|addressCity| City name|
|addressZip| Zip code|
|beds| Number of beds|
|baths| Number of baths|
|area| Sq. footage|
|latitude| Latitude|
|longitude| Longitude|
|variableData| String of description from Zillow|
|sgapt| For Sale by broker OR new construction (categ.)|
|zestimate| Zillow price estimate|
|brokerName| Broker (seller) name|

Features from the dataset have been removed for this analysis. The feature `unformattedPrice` is removed because it is the same as `price`. `address` and `addressStreet` are removed as they will not be useful in analysis or prediction because they're all unique and non-numeric. `addressState` only had the value "Texas". `isZillowOwned` is only FALSE as Zillow does not own any of the properties. `badgeInfo` is undefined and is NULL. `pgapt` is all "For Sale".
`info3string` is a link to a picture of a house.

```{r subsetted data, echo = F}
zillow_dat <- data %>%
  select(!c(unformattedPrice, address, addressStreet, addressState,
         isZillowOwned, badgeInfo, pgapt, info3String))

# head(zillow_dat)
```

```{r hist of price feat}
# hist(zillow_dat$price)
```

## EDA


### Checking Cities
We did not need the `addressState` column because it only had one value: Texas. We might assume that `addressCity` is only Austin and remove the feature, but is there another city in the dataset?

```{r check cities, echo = F, eval = F}
zillow_dat %>%
  # group by category
  group_by(addressCity) %>% 
  # Count the number of values assocuated with the category
  summarize(n = n()) %>%
  # Most listings first
  arrange(desc(n))
```
|City| Listings|
|----|---------|
|Austin|	786			
|Del Valle|	8			
|Pflugerville|	3			
|Lakeway|	1			
|Manor|	1			
|Terlingua|	1	

There are a few other cities represented in this dataset, so we will keep this column for now.

### Brand-New or Old house

It is also interesting to know which houses are being sold and under construction, or are for sale by a broker.

```{r new or old house (house type), echo = F, eval = F}
zillow_dat %>%
  # group by category
  group_by(sgapt) %>% 
  # Count the number of values assocuated with the category
  summarize(n = n()) %>%
  # Most listings first
  arrange(desc(n))
```
|Listing type| Listings|
|----------|---------|
|For Sale (Broker)|	755			
|New Construction|	45

Most of the houses for sale have already been built. Lets see if there is a price difference between the types of listings.

```{r listing type boxp, echo = F}
ggplot(zillow_dat, aes(x = sgapt, y = log(price), fill = sgapt)) +
  geom_boxplot() +
  labs(x = 'Listing Type', y = 'Log price',
       title = 'Log Price by Listing Type', fill = 'Listing type') +
  theme(plot.title = element_text(size=20))
```
Looks like buying a new house tends to cost more than one already for sale by a broker. Recall this is the log price.

So, let's look at the mean housing price of both listing types.

```{r listing type means, echo = F, eval = F}
zillow_dat %>%
  # group by category
  group_by(sgapt) %>% 
  # Count the number of values assocuated with the category
  summarize(mean(price))
```
|Listing type| Mean price|
|------------|-----------|
|For Sale (Broker)|	915,297.9			
|New Construction|	1,882,131.2

Looking at the mean price of the listing types is disheartening, not only do you have to be hip to live in Austin, but you have to be rich too? The mean prices of the two listing types are quite high, and we see from the boxplot that the data is skewed, so let's employ some resampling techniques to get an estimate on the true mean house prices.

It might be worth taking a closer look at the normality of the data.

```{r Histograms of Listing Types, echo = F, warning=F, fig.height=4, fig.width=5}
ggplot(zillow_dat, aes(x = log(price))) + 
  geom_histogram(binwidth = .1) +
  facet_wrap(~sgapt) +
  labs(title = 'Facet Histograms of Listing Types',
       x = 'Log price', y = 'Frequency')
```

We can already observe the skewedness of the Broker sale data from the boxplot, but it is hard to tell the shape of the "New Construction" data.

```{r New Construction Listing Hist, echo = F, fig.height=4, fig.width=5}
options(scipen=5) # Remove scientific notation from histogram
ggplot(subset(zillow_dat, sgapt == 'New Construction'),
       aes(x = price)) + geom_histogram(bins = 100) +
  labs(title = "Histogram of 'New Construction' Listing Prices",
       x = 'Price', y = 'Frequency')
  
```

So, there are quite a few houses over the million dollar price range... And the most expensive house is $11,250,800. These are probably for Elon Musk and his pals, though.

```{r Huge outlier listing in New Construction, echo = F, eval = F}
outlier_listing <- zillow_dat %>%
  filter(sgapt == 'New Construction') %>%
  arrange(desc(price)) %>%
  head(1)

# outlier_listing$price
```

Let's get a range estimate for the true mean house prices of both listing types by looking at the distribution of the bootstrap means.

```{r bootstrap mean and median Broker sale, echo = F, eval = T}
################## Bootstrap percentile technique #####################
subset_brokerSale <- zillow_dat %>%
  filter(sgapt == 'For Sale (Broker)') %>%
  select(price)

subset_brokerSale <- as.numeric(unlist(subset_brokerSale))
# class(subset_brokerSale)

obs.mean <- mean(subset_brokerSale)

# Length (number) of observations
# Dataframe will use nrow instead of length
n <- length(subset_brokerSale)

# Resample iterations
N <- 10^4

# Bootstrap mean container
boot.mean <- numeric(N)
# Bootstrap median container
boot.meds <- numeric(N)

# Bootstrap
for(i in 1:N)
{
  # Resample
  x <- sample(subset_brokerSale, n, replace = T)
  # Bootstrap means
  boot.mean[i] <- mean(x)
  # Bootstrap medians
  boot.meds[i] <- median(x)
}

# quantile(boot.mean, c(.025,.975))
# abs((mean(boot.mean) - 915297.9) / sd(boot.mean))
```

```{r hist of bootstrap means (brokerSale), echo = F, fig.height=4, fig.width=5}
graphics::hist(boot.mean, main = 'Bootstrap Distribution of Means \n For Broker Sales', xlab = 'Means')
```

We are 95% confident in saying that the true mean housing price in Austin is between (820715.3 1021421.0) and our bias percentage shows the sample mean may well represent the true mean, with the bias being only .6%. But we actually want people to move to Austin and the mean isn't the best descriptor in this case, so let's check out the bootstrap distribution of the median and see if we get a more reasonable price.


```{r boostrap percentile medians, echo = F, eval = F}
# Observed median
obs.med <- median(subset_brokerSale)
# 95% CI for true median
quantile(boot.meds, c(.025,.975))
# bias percentage on median
abs((mean(boot.meds) - 575000) / sd(boot.meds))
```

```{r hist of median (broker sale), echo = F, fig.height=4, fig.width=5}
graphics::hist(boot.meds, main = 'Bootstrap Distribution of Medians \n For Broker Sales', xlab = 'Medians')
```

We are 95% confident that (549000 599000) is one of the infinitely many 95% CIs that contain the true median house price. Our bias percentage is kind of high, at 10.89% biased. Anyway, around 500k - 600k for a house sounds a lot better than an average of 900k.

How does this compare to newly constructed houses? We saw that the average price for newly constructed housing units was approximately double that of ones for sale by a broker. Since the median looks like a better estimator, let's look at the median for newly constructed houses.

```{r bootstrapping medians on New Construction, echo = F}
################## Bootstrap percentile technique #####################
#######################################################################
subset_new_constr <- zillow_dat %>%
  filter(sgapt == 'New Construction') %>%
  select(price)

subset_new_constr <- as.numeric(unlist(subset_new_constr))
# class(subset_brokerSale)

obs.mean <- mean(subset_new_constr)

# Length (number) of observations
# Dataframe will use nrow instead of length
n <- length(subset_new_constr)

# Resample iterations
N <- 10^4

# Bootstrap mean container
boot.mean <- numeric(N)
# Bootstrap median container
boot.meds <- numeric(N)

# Bootstrap
for(i in 1:N)
{
  # Resample
  x <- sample(subset_new_constr, n, replace = T)
  # Bootstrap means
  # boot.mean[i] <- mean(x)
  # Bootstrap medians
  boot.meds[i] <- median(x)
}

# quantile(boot.mean, c(.025,.975))
# abs((mean(boot.mean) - obs.mean) / sd(boot.mean))

# graphics::hist(boot.mean, main = 'Bootstrap Distribution of Means \n For Broker Sales', xlab = 'Means')

# Observed median
obs.med <- median(subset_new_constr)
# 95% CI for true median
quantile(boot.meds, c(.025,.975))
# bias percentage on median
abs((mean(boot.meds) - obs.med) / sd(boot.meds))

graphics::hist(boot.meds, main = 'Bootstrap Distribution of Medians \n For Broker Sales', xlab = 'Medians')

```

The median is highly skewed, the interval estimate from the 95% CI has an enormous range, and the bias percentage is approximately 50% (.4786609), so it is not wise to trust population parameters for new houses coming from this sample.

Let's perform one more test on median housing price. The website realtor.com claims that, in November 2022, the median housing price was 600k USD [3]. We will use both listing types for the test. We will test the hypotheses:

$H_0: \tilde{\mu} <= 600,000$ versus 
$H_A: \tilde{\mu} > 600,000$

```{r hypothesis test of 600k median claim, echo = F}
# Observed median housing price
# for both listing types
obs.med <- median(zillow_dat$price)

# Sample size
n <- length(zillow_dat$price)

# Bootstrap iters
N <- 10^4

# Container
boot.med.full <- numeric(N)

for(i in 1:N)
{
  x <- sample(zillow_dat$price, n, replace = T)
  boot.med.full[i] <- median(x)
}

# p-value
# ( sum(boot.med.full >= obs.med) + 1) / (N + 1)

```

The p-value, being the probability that the statistic occurs randomly, is approximately 50.32%. Given the data, we fail to reject the null hypothesis at any small level of significance $\alpha$ and say that *there is not statistically significant evidence to contradict the claim that the median house price is more than 600k.*

So, we're fairly certain we could find some options under 600k. Austin seems like "the spot", though, so we may be willing to dish out some extra Quiche for a "spot at the spot".

### Top Brokers

Of course, life isn't all butterflies and candy: we have to make Quiche (cash), buy property, maintain the property... I digress.

Let's see which brokers we should call up to get a decent deal.

We think maybe we should go with the brands selling the most properties, as the buying process may be easier and the price may be lower. The Brokers with the most listings are:

```{r top 5 brokers, echo = F}
top_5_brokers <- zillow_dat %>%
  # group by category
  group_by(brokerName) %>% 
  # Count the number of values assocuated with the category
  summarize(n = n()) %>%
  # Most listings first
  arrange(desc(n)) %>%
  head(5)

# top_5_brokers$brokerName

# top_5_brokers
# sum(top_5_brokers$n)
```

| Broker| Listings|
|-------|---------|
|Realty Austin|	73			
|Keller Williams Realty|	67			
|Compass RE Texas, LLC|	60			
|Opendoor Brokerage, LLC|	32			
|Moreland Properties|	29	

Other than Keller Williams Realty, the other brokers are not huge names if you are not into the housing market. Let's explore the data relative to the top 5 brokers.

```{r subsetting brokers, echo = F}
by_top_5_brokers <- zillow_dat %>%
  filter(brokerName %in% top_5_brokers$brokerName)
```

```{r box top 5 brokers, echo = F}
ggplot(by_top_5_brokers, aes(x = brokerName, y = log(price), 
                             fill = brokerName)) +
  geom_boxplot() + 
  coord_flip() + 
  labs(title = 'Price Boxplots by Broker', y = 'Log price',
       x = 'Broker name')
```

It looks like Moreland Properties has the highest variation in house prices. Moreland is also has the fewest listings likely because they're so high priced.

```{r median house price by broker, echo = F}
zillow_dat %>%
  filter(brokerName %in% top_5_brokers$brokerName) %>%
  group_by(brokerName) %>%
  summarize(median = median(price))
```

Keller Williams, Austin Realty, and Compass RE look like they are doing similar business, though Compass has some huge outliers. If you are moving to the greater Austin area and your friend recommends using Moreland Properties, they probably think you are rich!



### References:

[1] https://www.austinchamber.com/economic-development/key-industries/corporate-headquarters

[2] https://www.kaggle.com/datasets/alexhuggler/austin-zillow-houses-for-sale

[3] https://www.realtor.com/realestateandhomes-search/Austin_TX/overview