---
title: "Analysis of Zillow Data from Austin, Texas"
author: "Will Curkan"
date: "2022-12-16"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo = F, warning=F, message=F, include=F}
library(dplyr, quietly = T)
```

## Introduction

Austin, a hilly city due south in the middle of Texas, is where all the young, hip Texans and out-of-staters are converging to live. Other than its famous 6-th street nightlife, the "weird" city is home to many "corporate and regional headquarters" from companies like Dell, IBM, Facebook, VISA and many more [1] - these are some great reasons to "kick up the roots" and mosey on over to Austin! 

## Objective Statement

The objective of this project is to make inferences about true statistic parameters in Austin, Texas, and to make predictions on housing prices and categories like which broker is listing the property.

## Data Acquisition and Explanation

```{r load data, echo = F}
set.seed(2022)
data <- read.csv('Zillow_Austin_11-16-22.csv')
#head(data)
#dim(data)
```
Alex Huggler on Kaggle.com provides Zillow Housing data for November 2022 on Kaggle.com [2]. The data has 800 unique observations and 20 features: one observation per unique housing unit and each feature as some trait of the house like its zip code or number of bedrooms.

### Feature Descriptions

The following features will be used in analysis of Austin housing prices:

|Feature|Description|
|-------|-----------|
|Price| price of housing|
|AddressCity| City name|
|AddressZip| Zip code|
|beds| Number of beds|
|baths| Number of baths|
|area| Sq. foot|
|latitude| Latitude|
|longitude| Longitude|
|variableData| String of description from Zillow|
|sgapt| For Sale by broker OR new construction (cat)|
|zestimate| Zillow how price estimate|
|brokerName| Who sells the house|

Features from the dataset have been removed for this analysis. The feature `unformattedPrice` is removed because it is the same as `price`. `address` and `addressStreet` are removed as they will not be useful in analysis or prediction because they're all unique and non-numeric. `addressState` only had the value "Texas". `isZillowOwned` is only FALSE as Zillow does not own any of the properties. `badgeInfo` is undefined and is NULL. `pgapt` is all "For Sale".
`info3string` is a link to a picture of a house.

```{r subsetted data, echo = F}
zillow_dat <- data %>%
  select(!c(unformattedPrice, address, addressStreet, addressState,
         isZillowOwned, badgeInfo, pgapt, info3String))

head(zillow_dat)
```

```{r hist of price feat}
# hist(zillow_dat$price)
```

## EDA


### Checking Cities
We did not need the `addressState` column because it only had one value: Texas. We might assume that `addressCity` is only Austin and remove the feature, but is there another city in the dataset?

```{r check cities, echo = F, eval = F}
zillow_dat %>%
  # group by category
  group_by(addressCity) %>% 
  # Count the number of values assocuated with the category
  summarize(n = n()) %>%
  # Most listings first
  arrange(desc(n))
```
|City| Listings|
|----|---------|
|Austin|	786			
|Del Valle|	8			
|Pflugerville|	3			
|Lakeway|	1			
|Manor|	1			
|Terlingua|	1	

There are a few other cities represented in this dataset, so we will keep this column for now.

### Top Brokers
First, let's look at the top 5 brokers with the most listings.

```{r top 5 brokers, echo = F}
top_5_brokers <- zillow_dat %>%
  # group by category
  group_by(brokerName) %>% 
  # Count the number of values assocuated with the category
  summarize(n = n()) %>%
  # Most listings first
  arrange(desc(n)) %>%
  head(5)

# top_5_brokers
# sum(top_5_brokers$n)
```
| Broker| Listings|
|-------|---------|
|Realty Austin|	73			
|Keller Williams Realty|	67			
|Compass RE Texas, LLC|	60			
|Opendoor Brokerage, LLC|	32			
|Moreland Properties|	29	

Other than Keller Williams Realty, the other brokers are not huge names if you are not into the housing market. Let's explore the data relative to the top 5 brokers.

```{r subsetting brokers, echo = F}
by_top_5_brokers <- zillow_dat %>%
  filter(brokerName %in% top_5_brokers$brokerName)
```



### References:

[1] https://www.austinchamber.com/economic-development/key-industries/corporate-headquarters

[2] https://www.kaggle.com/datasets/alexhuggler/austin-zillow-houses-for-sale