---
title: "Analysis of Zillow Data from Austin, Texas"
author: "Will Curkan"
date: "2022-12-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Austin, a hilly city due south in the middle of Texas, is where all the young, hip Texans and out-of-staters are converging to live. Other than its famous 6-th street nightlife, the "weird" city is home to many "corporate and regional headquarters" from companies like Dell, IBM, Facebook, VISA and many more [1] - these are some great reasons to "kick up the roots" and mosey on over to Austin! 

## Data Acquisition and Explanation

Alex Huggler on Kaggle.com provides Zillow Housing data for November 2022 on Kaggle.com [2]. 

```{r echo = F}
zillow.dat <- read.csv('Zillow_Austin_11-16-22.csv')
```



### References:

[1] https://www.austinchamber.com/economic-development/key-industries/corporate-headquarters

[2] https://www.kaggle.com/datasets/alexhuggler/austin-zillow-houses-for-sale